RNAseq aCRY pCRY data dive (separate)

### 2024-04-26 2_A_dive_aCRY_pCRY (separately)/2_A_dive_aCRY_pCRY_sep.Rmd
figure fix

# Load System
## -R_libraries
```{r r_libraries}
library(patchwork)
library(PCAtools)
library(stringr)
library(R.utils)
library(RColorBrewer)
library(sessioninfo)
library(data.table)
library(plyr)
library(tidyverse)
library(tximeta)
library(tximport)
library(curl)
library(AnnotationHub)
library(DESeq2)
library(EnhancedVolcano)
library(pheatmap)
library(writexl)
library(biomaRt)
library(ape)
library(kableExtra)
library(knitr)

@@

## -R_folders
```{r r_folders}

@@

# 1. Load aCRY & pCRY
## - Tximeta files
```{r get_tximeta_files}

@@

## Run Deseq2
```{r dds_acry_pcry}

@@

##########################################\n
# filter all rows with rowsum = 0 ####\n
par(mfrow=c(1,2))
hist(log(counts(dds)), breaks=100, ylim = c(0,11000), xlim = c(0,10))
hist(counts(dds), breaks=1000000, ylim = c(0,100000), xlim = c(0,10))
# -> dip at log(counts(dds)=2-3)
par(mfrow=c(1,1))

@@

plotDispEsts(dds)
resultsNames(dds)
DESeq2::plotMA(dds)

plotCounts(dds, gene = "Cre01.g000150", intgroup = "condition", col=colData(dds)$condition)

@@

```{r pre_trans_fig, figures-side, fig.show="hold", out.width="25%"}
# load(file = paste(dir,"rlog_list.RDS",sep="/"))
msdp1 <- lapply(lapply(dds_list,assay,normalized =TRUE),meanSdPlot)
msdp2 <- lapply(lapply(vsd_list,assay,normalized =TRUE),meanSdPlot)
msdp3 <- lapply(lapply(ntd_list,assay,normalized =TRUE),meanSdPlot)
msdp4 <- lapply(lapply(rld_list,assay,normalized =TRUE),meanSdPlot)

@@

### - Check sample distance
```{r pre_sample_dist, fig.height=6, out.width="100%", eval=TRUE, include=TRUE, echo=FALSE}
vsd <- vsd_list[[1]]
for (vsd in vsd_list){
sampleDists <- dist(t(assay(vsd)))

@@

### - Perform principal component analysis
```{r pca, warning = FALSE, fig.show="hold", out.width="80%", fig.height=7, eval=TRUE, include=TRUE, echo=FALSE}
# transform data
# load(file=paste(data,"deseq2.dds", sep="/"))

@@

# 3. Results
## Colors
```{r}
# colours
Greys <- brewer.pal(9, name="Greys")[c(7,5)]
Paired <- brewer.pal(12, name="Paired")[c(2,1,6,5)]
colours <- c(Greys,Paired)
group.colors <- colours
group.colors <- c("#525252", "#969696", "#1F78B4", "#A6CEE3", "#E31A1C", "#FB9A99")
# group.colors <- c("grey10","grey40","grey80","royalblue4","royalblue2", "lightblue2",  "darkred","brown2","salmon1")

@@

## Plot Counts
```{r plot_counts}
dds <- dds_list[[1]]
for (dds in dds_list){
# TF genes

@@

print(gcounts_tf1 / gcounts_phot1 & plot_annotation(title = colData(dds)$experiment[1]))

@@

# ---------------------------
# COPY & PASTE from pCRY
## Make results
```{r, eval=FALSE, include=FALSE, echo=FALSE}
outdir <- "C:/Users/kelterbs/OneDrive - Charité - Universitätsmedizin Berlin/NGS (RNA-Seq & CRISPR)/Chlamy RNA-Seq P3044"
setwd(outdir)

@@

## Colours
```{r, eval=FALSE}
Greys <- brewer.pal(9, name="Greys")[c(7,5)]
Paired <- brewer.pal(12, name="Paired")[c(2,1,6,5)]
colours <- c(Greys,Paired)
```

## PCA (Panel B)
```{r, eval=FALSE}
library("PCAtools")
library(ggalt)

@@

## Counts
```{r, eval=FALSE}

@@

@@

## Heatmap

>>>>

# Load System
## -R_libraries
```{r r_libraries, imnclude=FALSE}
library(patchwork)
library(PCAtools)
library(stringr)
library(R.utils)
library(RColorBrewer)
library(sessioninfo)
library(data.table)
library(plyr)
library(tidyverse)
library(tximeta)
library(tximport)
library(curl)
library(AnnotationHub)
library(DESeq2)
library(EnhancedVolcano)
library(pheatmap)
library(writexl)
library(biomaRt)
library(ape)
library(kableExtra)
library(knitr)

@@

## -R_folders
```{r r_folders, include = FALSE}

@@

# 1. Load aCRY & pCRY
## - Tximeta files
```{r get_tximeta_files, include=FALSE}

@@

## Run Deseq2
```{r dds_acry_pcry, echo=FALSE}

@@

##########################################\n
# filter all rows with rowsum = 0 ####\n
par(mfrow=c(1,2))
hist(log(counts(dds)), breaks=100, ylim = c(0,11000), xlim = c(0,10), main = colData(dds)$experiment[1]) 
hist(counts(dds), breaks=1000000, ylim = c(0,100000), xlim = c(0,10))
# -> dip at log(counts(dds)=2-3)
par(mfrow=c(1,1))

@@

plotDispEsts(dds, main = colData(dds)$experiment[1])
resultsNames(dds)
DESeq2::plotMA(dds, main=colData(dds)$experiment[1])

ROC40 <- "Cre06.g275350"
plotCounts(dds, gene = ROC40, intgroup = "condition", col=colData(dds)$treatment, main = paste("ROC40 in Δ",colData(dds)$experiment[1],sep = ""))

@@

```{r pre_trans_fig, figures-side, fig.show="hold", out.width="50%"}
# load(file = paste(dir,"rlog_list.RDS",sep="/"))
msdp1 <- lapply(lapply(dds_list,assay,normalized =TRUE),meanSdPlot)
msdp2 <- lapply(lapply(vsd_list,assay,normalized =TRUE),meanSdPlot)
msdp3 <- lapply(lapply(ntd_list,assay,normalized =TRUE),meanSdPlot)
msdp4 <- lapply(lapply(rld_list,assay,normalized =TRUE),meanSdPlot)

@@

### - Check sample distance
```{r pre_sample_dist, fig.height=4, out.width="100%", eval=TRUE, include=TRUE, echo=FALSE}
vsd <- vsd_list[[1]]
for (vsd in vsd_list){
sampleDists <- dist(t(assay(vsd)))

@@

### - Perform principal component analysis
```{r pca, warning = FALSE, fig.show="hold", out.width="80%", fig.height=5, eval=TRUE, include=TRUE, echo=FALSE}
# transform data
# load(file=paste(data,"deseq2.dds", sep="/"))

@@

# 3. Results
## Colors
```{r colors, echo=FALSE}
# colours
Greys <- brewer.pal(9, name="Greys")[c(7,5)]
Paired <- brewer.pal(12, name="Paired")[c(2,1,6,5)]
colours <- c(Greys,Paired)
group.colors <- colours
group.colors <- c("#525252", "#969696", "#1F78B4", "#A6CEE3", "#E31A1C", "#FB9A99")
# group.colors <- c("grey10","grey40","grey80","royalblue4","royalblue2", "lightblue2",  "darkred","brown2","salmon1")

@@

## Plot Counts
```{r plot_counts, echo=FALSE}
dds <- dds_list[[1]]
for (dds in dds_list){
# TF genes

@@

# print(gcounts_tf1 / gcounts_phot1 & plot_annotation(title = colData(dds)$experiment[1]))

# single gene
    d <- plotCounts(dds, gene=ROC40, intgroup=c("condition","experiment","genotype","treatment","clientName"), main=s,returnData=TRUE)

gcounts <- ggplot(d, aes(x = treatment, y = count, fill=condition, color=condition)) +
    geom_boxplot(color="black") +
    geom_point(shape=21,color="black",aes(fill=condition),position=position_dodge(width=0.75), alpha=1) +
  geom_text_repel(aes(label=clientName)) +
    scale_fill_manual(values=group.colors) +
    scale_color_manual(values=group.colors) +
    scale_y_continuous(trans = "log2") +
    labs(title = paste("ROC40"," in Δ",colData(dds)$experiment[1],sep = ""))
gcounts %>% print()

@@

#### ---------------------------
#### COPY & PASTE from pCRY
#### Make results
```{r, eval=FALSE, include=FALSE, echo=FALSE}
outdir <- "C:/Users/kelterbs/OneDrive - Charité - Universitätsmedizin Berlin/NGS (RNA-Seq & CRISPR)/Chlamy RNA-Seq P3044"
setwd(outdir)

@@

### Colours
```{r, eval=FALSE}
Greys <- brewer.pal(9, name="Greys")[c(7,5)]
Paired <- brewer.pal(12, name="Paired")[c(2,1,6,5)]
colours <- c(Greys,Paired)
```

### PCA (Panel B)
```{r, eval=FALSE}
library("PCAtools")
library(ggalt)

@@

### Counts
```{r, eval=FALSE}

@@

### Volcanos
### dark
```{r, eval=FALSE}

@@

### Heatmap
